---
title: "*Spatial Statistics Lab 4*"
author: '**Onyedikachi J Okeke**'
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **0.0 To Load the library**
Package rgda and ‘maptools’ was removed from the CRAN repository.
Formerly available versions can be obtained from the archive.
Archived on 2023-10-16 at the request of the maintainer.
So Consider using 'sf' or 'terra' instead

```{r,comment="",echo=TRUE,eval=FALSE}
library(sp)
library(spatstat)
library(sf)
library(spatstat.geom)
library(ctv)
library(terra)
library(spdep)
library(terra)
library(RColorBrewer)
library(classInt)
library(epitools)
library(DCluster)
library(rgdal)
library(maptools)

```
Library 
```{r,echo=F,results='Show', fig.show='hold',eval=T}
library(sp)
library(spatstat)
library(sf)
library(spatstat.geom)
library(ctv)
library(terra)
library(spdep)
library(terra)
library(RColorBrewer)
library(classInt)
library(epitools)
library(DCluster)
library(rgdal)
library(maptools)

```

### **Loading city_limits_km and StLouisCrime2014**

```{r,comment="",echo=TRUE,eval=FALSE}
library(rgdal)
library(spatstat)
library(sp)
library(sf)
library(spatstat.geom)
library(maptools)

# Read the shapefile
S <- readOGR("C:/Spatial Statistics Labwork/Lab4Data/city_limits_km.shp")

# Convert the Spatial object to SpatialPolygons
SP <- as(S, "SpatialPolygons")

SP <- as(S, "SpatialPolygons")

W <- as(SP, "owin")

plot(W)

xy <- read.table("C:/Spatial Statistics Labwork/Lab4Data/StLouisCrime2014.txt", header=T, sep="\t")
attach(xy)
pp <- ppp(X, Y, window=W, marks=CRIME)

plot(pp)

gun <- pp[CRIME=="DISORDERLY"]
rob <- pp[CRIME=="BURGLARY"]
hit <- pp[CRIME=="HITANDRUN"]

plot(density(gun))
contour(density(gun), add=T)
plot(gun, add=T)

#Kernel density visualization is performed in spatstat using the density() function

plot(density(gun, 0.25))

d250 <- density(gun, 0.25)
plot(d250)
contour(d250, add=T)
plot(gun, add=T)

plot(gun, add=T)
r <- bw.diggle(gun)

r

```
##### Result
```{r,echo=F,results='Show', fig.show='hold',eval=T}
library(rgdal)
library(spatstat)
library(sp)
library(sf)
library(spatstat.geom)
library(maptools)

# Read the shapefile
S <- readOGR("C:/Spatial Statistics Labwork/Lab4Data/city_limits_km.shp")

# Convert the Spatial object to SpatialPolygons
SP <- as(S, "SpatialPolygons")

SP <- as(S, "SpatialPolygons")

W <- as(SP, "owin")

plot(W)

xy <- read.table("C:/Spatial Statistics Labwork/Lab4Data/StLouisCrime2014.txt", header=T, sep="\t")
attach(xy)
pp <- ppp(X, Y, window=W, marks=CRIME)

plot(pp)

gun <- pp[CRIME=="DISORDERLY"]
rob <- pp[CRIME=="BURGLARY"]
hit <- pp[CRIME=="HITANDRUN"]

plot(density(gun))
contour(density(gun), add=T)
plot(gun, add=T)

#Kernel density visualization is performed in spatstat using the density() function

plot(density(gun, 0.25))

d250 <- density(gun, 0.25)
plot(d250)
contour(d250, add=T)
plot(gun, add=T)

plot(gun, add=T)
r <- bw.diggle(gun)

r
```

Q4. Create density maps (in R) of the gun homicide data, experimenting with different kernel density bandwidths. Provide a commentary discussing the most suitable bandwidth choice for this analysis visualization method. (15 points)


## **convert it to a spatstat point pattern object, with the different crime types as an identifying mark**

```{r,comment="",echo=TRUE,eval=FALSE}
nnd <- nndist.ppp(gun)
hist(nnd)
summary(nnd)
mnnd <- mean(nnd)
exp_nnd <- 0.5 / sqrt(gun$n / area.owin(W))

print (mnnd / exp_nnd)

```
#### Result
```{r,echo=F,results='Show', fig.show='hold',eval=T}
nnd <- nndist.ppp(gun)
hist(nnd)
summary(nnd)
mnnd <- mean(nnd)
exp_nnd <- 0.5 / sqrt(gun$n / area.owin(W))

print (mnnd / exp_nnd)
```

Q5. Is this pattern clustered? Or evenly-spaced?


## **The Quadrat analysis**

```{r,comment="",echo=TRUE,eval=FALSE}
par(mfrow=c(1,1))
q <- quadratcount(hit, 4, 8)
plot(q)
plot(hit, add=T)
quadrat.test(hit, 4, 8)
quadrat.test(hit, 4, 8, alternative="clustered")
quadrat.test(hit, 4, 8, alternative="regular")

```
#### Result
```{r,echo=F,results='Show', fig.show='hold',eval=T}
par(mfrow=c(1,1))
q <- quadratcount(hit, 4, 8)
plot(q)
plot(hit, add=T)
quadrat.test(hit, 4, 8)
quadrat.test(hit, 4, 8, alternative="clustered")
quadrat.test(hit, 4, 8, alternative="regular")

```

Q6. Take a screenshot of the plotted figure. Is it a clustered, regular, or random pattern?

## **The distance-based functions: G, F, K (and its relative L) and the more recent pair correlation function.**

```{r,comment="",echo=TRUE,eval=FALSE}

library(spatstat) 

g_gun <- Gest(gun)
f_gun <- Fest(gun)
k_gun <- Kest(gun)
l_gun <- Lest(gun)
pcf_gun <- pcf(gun)

par(mfrow = c(1, 1))

plot(g_gun)
plot(f_gun)
plot(k_gun)
plot(l_gun)
plot(pcf_gun)

```
#### Result
```{r,echo=F,results='Show', fig.show='hold',eval=T}

g_gun <- Gest(gun)
f_gun <- Fest(gun)
k_gun <- Kest(gun)
l_gun <- Lest(gun)
pcf_gun <- pcf(gun)

par(mfrow = c(1, 1))
plot(g_gun)
plot(f_gun)
plot(k_gun)
plot(l_gun)
plot(pcf_gun)

```
